The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\ with the Owens Valley \revision{Radio Observatory} Long Wavelength Array

### 2018-03-06 paper.tex
mark revisions from the submitted version in red

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}\PackageWarning{TODO:}{#1!}}

@@

\title{
    The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\
    with the Owens Valley Radio Observatory Long Wavelength Array
}

@@

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21-cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \citet{2017MNRAS.464.4995M} as
\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.6}\,.
\end{equation}
Therefore, experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission is spectrally smooth. The low-frequency radio sky is composed of several components:
galactic synchrotron emission, supernova remnants, radio galaxies, free-free emission and absorption
from \ion{H}{2} regions, and a confusing background of radio sources.  Ideally, a foreground removal
strategy should be informed by the measured spatial structure and frequency spectrum of all
foreground components. For instance, CMB experiments typically construct several maps at several
frequencies to enable component separation.  At low frequencies, this possibility is limited by the
availability of suitable high-fidelity sky maps on angular scales ranging from tens of degrees to
arcminutes.

@@

As a demonstration of this technique we apply Tikhonov regularized $m$-mode analysis imaging to the
Owens Valley Radio Observatory Long Wavelength Array (OVRO-LWA), and generate a series of new
low-frequency maps of the sky between 36.528~MHz and 73.152~MHz.  These maps capture the full sky
visible from the Owens Valley Radio Observatory (OVRO) with angular resolution of $\sim 15$~arcmin.
These new maps complement the existing full-sky maps at these frequencies with greatly improved
angular resolution.

@@

The structure of this paper is as follows. In \S\ref{sec:imaging}, we present Tikhonov-regularized
$m$-mode analysis imaging, a new imaging technique that allows us to image the entire visible sky in
one coherent synthesis imaging step with exact widefield corrections. In \S\ref{sec:observations} we
describe our observations with the OVRO-LWA. In \S\ref{sec:results} we present the sky maps and
compare these maps against other low-frequency sky maps.  In \S\ref{sec:error} we discuss some of
the sources of error present in the maps, and finally in \S\ref{sec:conclusion} we present our
conclusions.

@@

The key advantage of $m$-mode analysis is the block-diagonal structure of
Equation~\ref{eq:m-mode-matrix-equation}. The computational complexity of many common matrix
operations (eg. solving a linear system of equations) is $\mathcal{O}(N^3)$, where $N$ is the linear
size of the matrix.  By splitting the equation into $M$ independent blocks, the number of floating
point operations required to solve the linear system of equations is now $\mathcal{O}(N^3M^{-2})$,
because each block can be manipulated independently of the other blocks. This computational savings
is what makes this matrix algebra approach to interferometric imaging feasible. For the dataset
presented in this paper, computing the elements of the transfer matrix takes $\sim$10 hours per
frequency channel on a 10-node cluster, but once the matrix has been computed the imaging process
described in \S\ref{sec:mmode-imaging} takes $\sim$10 minutes, and the deconvolution process
described in \S\ref{sec:clean} was allowed to run for $\sim$10 hours.

@@

\subsection{The Owens Valley Radio Observatory Long Wavelength Array}

@@

The Owens Valley Radio Observatory Long Wavelength Array (OVRO-LWA) is a 288-element interferometer
located at the Owens Valley Radio Observatory near Big Pine, California \citep{hallinan_2017}.  The
OVRO-LWA is a low-frequency instrument with instantaneous bandwidth covering 27~MHz to 85~MHz and
with 24~kHz channelization.  Each antenna stand hosts two perpendicular broadband dipoles so that
there are $288\times2$ signal paths in total. These signal paths feed into the 512-input LEDA
correlator \citep{2015JAI.....450003K}, which allows the OVRO-LWA to capture the entire visible
hemisphere in a single snapshot image.

@@

The noise in each map is empirically measured using jackknife resampling. The dataset is first split
into even- and odd-numbered integrations. These two groups are then imaged and CLEANed
independently, before being compared against the maps constructed from all of the available data
using the jackknife standard error estimator. This estimate of the standard error includes all
sources of error that operate on $\sim13$ second timescales (the integration time) such as thermal
noise and rapid ionospheric fluctuations, but does not account for more slowly varying effects (for
example, sidereal variation in the system temperature or day-night fluctuations in the ionosphere).
These noise calculations are summarized in Table~\ref{tab:summary}.  VLSSr source counts
\citep{2014MNRAS.440..327L} suggest that the confusion limit at 74 MHz and 15 arcmin angular
resolution is $\sim 1000\times(\nu/74\,{\rm MHz})^{-0.7}$ mJy.  Each channel map achieves thermal
noise $<900$ mJy and therefore each map is likely at or near the confusion limit.

>>>>

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}\PackageWarning{TODO:}{#1!}}
\newcommand{\revision}[1]{\textcolor{red}{\textbf{#1}}}

@@

\title{
    The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\
    with the Owens Valley \revision{Radio Observatory} Long Wavelength Array
}

@@

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21-cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \revision{\citet{2017MNRAS.464.4995M}} as
\revision{\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.6}\,.
\end{equation}}
Therefore, experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission is spectrally smooth. The low-frequency radio sky is composed of several components:
galactic synchrotron emission, supernova remnants, radio galaxies, free-free emission and absorption
from \ion{H}{2} regions, and a confusing background of radio sources.  Ideally, a foreground removal
strategy should be informed by the measured spatial structure and frequency spectrum of all
foreground components. For instance, CMB experiments typically construct several maps at several
frequencies to enable component separation.  At low frequencies, this possibility is limited by the
availability of suitable high-fidelity sky maps on angular scales ranging from tens of degrees to
arcminutes.

@@

As a demonstration of this technique we apply Tikhonov regularized $m$-mode analysis imaging to the
Owens Valley \revision{Radio Observatory} Long Wavelength Array (OVRO-LWA), and generate a series of new
low-frequency maps of the sky between 36.528~MHz and 73.152~MHz.  These maps capture the full sky
visible from the Owens Valley Radio Observatory (OVRO) with angular resolution of $\sim 15$~arcmin.
These new maps complement the existing full-sky maps at these frequencies with greatly improved
angular resolution.

@@

The structure of this paper is as follows. In \S\ref{sec:imaging}, we present Tikhonov-regularized
$m$-mode analysis imaging, a new imaging technique that allows us to image the entire visible sky in
one coherent synthesis imaging step with exact widefield corrections. In \S\ref{sec:observations} we
describe our observations with the \revision{OVRO-LWA}. In \S\ref{sec:results} we present the sky
maps and compare these maps against other low-frequency sky maps.  In \S\ref{sec:error} we discuss
some of the sources of error present in the maps, and finally in \S\ref{sec:conclusion} we present
our conclusions.

@@

The key advantage of $m$-mode analysis is the block-diagonal structure of
Equation~\ref{eq:m-mode-matrix-equation}. The computational complexity of many common matrix
operations (eg. solving a linear system of equations) is $\mathcal{O}(N^3)$, \revision{where $N$ is
the linear size of the matrix.  By splitting the equation into $M$ independent blocks, the number of
floating point operations required to solve the linear system of equations is now
$\mathcal{O}(N^3M^{-2})$, because each block can be manipulated independently of the other blocks.
This computational savings is what makes this matrix algebra approach to interferometric imaging
feasible. For the dataset presented in this paper, computing the elements of the transfer matrix
takes $\sim$10 hours per frequency channel on a 10-node cluster, but once the matrix has been
computed the imaging process described in \S\ref{sec:mmode-imaging} takes $\sim$10 minutes, and the
deconvolution process described in \S\ref{sec:clean} was allowed to run for $\sim$10 hours.}

@@

\subsection{The Owens Valley \revision{Radio Observatory} Long Wavelength Array}

@@

The Owens Valley \revision{Radio Observatory} Long Wavelength Array (OVRO-LWA) is a 288-element
interferometer located at the Owens Valley Radio Observatory near Big Pine, California
\citep{hallinan_2017}.  The OVRO-LWA is a low-frequency instrument with instantaneous bandwidth
covering 27~MHz to 85~MHz and with 24~kHz channelization.  Each antenna stand hosts two
perpendicular broadband dipoles so that there are $288\times2$ signal paths in total. These signal
paths feed into the 512-input LEDA correlator \citep{2015JAI.....450003K}, which allows the OVRO-LWA
to capture the entire visible hemisphere in a single snapshot image.

@@

The noise in each map is empirically measured using jackknife resampling. The dataset is first split
into even- and odd-numbered integrations. These two groups are then imaged and CLEANed
independently, before being compared against the maps constructed from all of the available data
using the jackknife \revision{standard error} estimator. This estimate of the \revision{standard
error} includes all sources of error that operate on $\sim13$ second timescales (the integration
time) such as thermal noise and rapid ionospheric fluctuations, but does not account for more slowly
varying effects (for example, sidereal variation in the system temperature or day-night fluctuations
in the ionosphere).  These noise calculations are summarized in Table~\ref{tab:summary}.  VLSSr
source counts \citep{2014MNRAS.440..327L} suggest that the confusion limit at 74 MHz and 15 arcmin
angular resolution is $\sim 1000\times(\nu/74\,{\rm MHz})^{-0.7}$ mJy.  Each channel map achieves
thermal noise $<900$ mJy and therefore each map is likely at or near the confusion limit.
